---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading, echo = TRUE}
data <- read.csv("activity.csv", header = TRUE, sep = ",")
```

 
### Converting Date to date format and Interval to factor
```{r convert, echo = TRUE}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$interval <- factor(data$interval)
```


## What is mean total number of steps taken per day?

### First find the sum of steps taken every day and plot the graph

```{r , echo = TRUE}
sm <- tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(sm, xlab = "no of steps taken per Day", main = "Histogram Of Steps taken per Day")
```
### For Mean and Median of No of Steps taken per day

```{r ,echo = TRUE}
sm_mean <- round(mean(sm))
sm_median <- round(median(sm))
print(c("The mean is " , sm_mean))
print(c("The median is ", sm_median))
```


## What is the average daily activity pattern?

### We have to calculate the mean of steps taken daily 

```{r interval , echo = TRUE}
mn <- tapply(data$steps, data$interval, mean, na.rm =TRUE)
plot(mn ~ unique(data$interval), typle = "l", xlab = "5 Min intervals")
```
### For Max no. of steps taken on a particular day. 
#### the following code will show the day on which maximum steps were take and how many.
```{r , echo = TRUE}
mn[which.max(mn)]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
