---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading, echo = TRUE}
data <- read.csv("activity.csv", header = TRUE, sep = ",")
```

 
### Converting Date to date format and Interval to factor
```{r convert, echo = TRUE}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$interval <- factor(data$interval)
```


## What is mean total number of steps taken per day?

### First find the sum of steps taken every day and plot the graph

```{r , echo = TRUE}
sm <- tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(sm, xlab = "no of steps taken per Day", main = "Histogram Of Steps taken per Day")
```
### For Mean and Median of No of Steps taken per day

```{r ,echo = TRUE}
sm_mean <- round(mean(sm))
sm_median <- round(median(sm))
print(c("The mean is " , sm_mean))
print(c("The median is ", sm_median))
```


## What is the average daily activity pattern?

### We have to calculate the mean of steps taken daily 

```{r interval , echo = TRUE}
mn <- tapply(data$steps, data$interval, mean, na.rm =TRUE)
plot(mn ~ unique(data$interval), typle = "l", xlab = "5 Min intervals")
```
### For Max no. of steps taken on a particular day. 


#### The following code will show the day on which maximum steps were take and how many.
```{r , echo = TRUE}
mn[which.max(mn)]
```

## Imputing missing values

### For the missing values we will replace it with the mean of that particular day.
#### We will create a new data set newData.

```{r , echo =TRUE}
newData <- data
  for(i in 1:nrow(data))
  { 
    if(is.na(data$steps[i]))
    {
    newData$steps[i] <- mn[[as.character(data[i,"interval"])]]
    }
  }
```
  
#### Now We will draw a histogram of steps taken with the new values 


```{r , echo =TRUE}
    sm2 <- tapply(newData$steps, newData$date, sum, na.rm = TRUE)
    hist(sm2 , xlab ="Total Steps taken per Day", main = "Histogram of total steps taken per Day")
```


#### We will calculate the mean and median of the new data set with no misssing values

  ```{r ,echo = TRUE}
  sm2_mean <- round(mean(sm2))
  sm2_median <- round(median(sm2))
  print(c("The new mean is " , sm2_mean))
  print(c("The new median is ", sm2_median))
  ```  
    

## Are there differences in activity patterns between weekdays and weekends?


### We'll add new column to the data set containing weekday or weekend

```{r , echo = TRUE}
newData$weekday <- c("Weekday")
newData[weekdays(as.Date(newData[,2])) %in% c("Saturday", "Sunday", "saturday", "sunday"),][4] <- c("Weekend")
table(newData$weekday == "Weekend")

newData$weekday <- factor(newData$weekday)

newData_weekday <- subset(newData, newData$weekday == "Weekday")
newData_weekend <- subset(newData, newData$weekday == "Weekend")
 mean_weekday <- tapply(newData_weekday$steps, newData_weekday$interval, mean)
 mean_weekend <- tapply(newData_weekend$steps, newData_weekend$interval, mean)
 ```
 
### Now we'll plot the graph

```{r ,echo =TRUE}
wd <- data.frame(interval = unique(newData_weekday$interval), avg = as.numeric(mean_weekday), day = rep("Weekday", length(mean_weekday)))

wnd <- data.frame(interval = unique(newData_weekend$interval), avg = as.numeric(mean_weekend), day = rep("Weekend", length(mean_weekend)))

final <- rbind(wd,wnd)
library(lattice)

 xyplot(avg ~ interval | day, data = final , layout = c(1,2), type = "l", ylab = "No. of Steps")
 ```
 
